---
title: "Becas"
author: "Lluís Revilla Sancho"
date: "4/27/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploring the I+D+i in Spain

There is lot of discussion, here I will analyze some data:

From the documen the more important ones are the 

> Anexo I: Relación de proyectos de investigación con propuesta de financiación
> Anexo II: Propuesta de financiación de cada proyecto
> Anexo III: ...condiciones de la ayuda...
> Anexo IV: Relación de proyectos a los que no se propone financiación
> Anexo V: ITI CÁDIZ


For reference here are some numbers
https://twitter.com/AgEInves/status/1121693189897723990

But some think that they are wrong:

https://twitter.com/BenLehner/status/1121812355019616256

So I went to the original data to see what happens

Those projects with and without money.


## Original data

The data mentioned in the tweet is [this one](https://sede.micinn.gob.es/stfls/eSede/Ficheros/2019/Propuesta_Resolucion_Provisional_Proyectos_Generacion_Conocimiento_2018_fda.pdf), however, my PI recieved the email with a link to another [pdf](http://www.ciencia.gob.es/stfls/eSede/Ficheros/2019/PRP_Retos_Investigacion_2018.pdf). I'll use the later
We need to extract the tables to be able to analyze the content.

## Obtaining the data

First we need to convert from pdf to text. A quick search pointed me to an [entry](https://medium.com/@CharlesBordet/how-to-extract-and-clean-data-from-pdf-files-in-r-da11964e252e) which recommends [`pdftools`](https://cran.r-project.org/package=pdftools)

```{r pdftools}
library("pdftools")
pdf <- "http://www.ciencia.gob.es/stfls/eSede/Ficheros/2019/PRP_Retos_Investigacion_2018.pdf"
text <- pdf_text(pdf)
document <- sapply(text, function(page) {
  lines <- strsplit(page, "\\n\\s+")
  lapply(lines, trimws)
}, USE.NAMES = FALSE)
```

By manually inspecting the file we find that the relevant "anexos" are in these pages:
```{r positions}
anex1 <- 4:118
anex2 <- 119:182
anex4 <- 190:309
interesting <- list(anex1, anex2, anex4)
names(interesting) <- c("anex1", "anex2", "anex4")
```
I restrict to those in table format and for the whole country.

Now we need to convert them to tables, by first removing the header and the notes at the end of the tables.
```{r trimming}
trim_page <- function(x) {
  header <- c(1:3)
  notes <- seq(from = length(x) -1 , to = length(x)) 
  x[-c(header, notes)]
}

anexes <- lapply(interesting, function(x, document) {
  lapply(document[x], trim_page)
}, document = document)

```
Now we want to split by field and we use this opportunity to correct some common mishandle of names of the research center.

We see that often the long names of the 5th field are in the previous line. However it is not correelcty parsed "AGENCIA ESTATAL CONSEJO SUPERIOR DE" is for the 4th field usually and the rest for the 5th.

```{r split_fields}
anexes_field <- lapply(anexes, function(x){
  lapply(x, function(pages) {
    y <- strsplit(pages, "\\s{2, }")
    zy <- y
    zy <- unlist(zy)
    y[lengths(y) == 1] <- sapply(y[lengths(y) == 1], function(x){
      # browser()
     if (grepl("AGENCIA ESTATAL CONSEJO SUPERIOR DE", x)) {
       z <- gsub("AGENCIA ESTATAL CONSEJO SUPERIOR DE ", "", x)
       c(z, "AGENCIA ESTATAL CONSEJO SUPERIOR DE")
     } else {
       x
     }
    })
    y
  })
})
```

# Anex 1

```{r}
for (anex in names(anexes_field)) {
  for (page in anexes_field[[anex]]) {
    fields_page <- lengths(page)
    size_table <- switch(anex, 
                         "anex1" = 9, 
                         "anex2" = 6,
                         "anex4" = 7)
    entries <- which(fields_page == size_table)
    n_lines <- length(fields_page)
    for (line in seq_len(n_lines)) {
      if ((fields_page == size_table)[line]) {
        previous_entry <- which.max(entries[-seq(from = 1, to = line-1)])
        interval <- seq(from = previous_entry, to = line -1)
        field4 <- paste(sapply(page[interval], "[", 1))
        field5 <- paste(sapply(page[interval], "[", -1))
        page[[line]][4] <- paste(field4, page[[line]][4])
        page[[line]][5] <- paste(field4, page[[line]][5])
      }
    }
    anexes_field[[anex]] <- page
  }
}

anext <- sapply(anexes_field[[1]], function(page){
  page[lengths(page) == 9]
})
anext2 <- lapply(anext[lengths(anext) > 1], function(page){
  t(simplify2array(page))
})
t2 <- do.call(rbind, anext2)
t2 <- as.data.frame(t2)
colnames(t2) <- c("reference", "area", "subarea", "entity", "center", "NIF", 
                  "CA", "years", "predoc")
fact2_number <- function(f){
   as.numeric(levels(f))[f]
}
t2$predoc <- fact2_number(t2$predoc)
t2$years <- fact2_number(t2$years)
```

```{r relevel}
library("forcats")
t2$CA <- fct_relabel(t2$CA,  ~ gsub("^A$", "EXTREMADURA", .))
t2$CA <- fct_relabel(t2$CA,  ~ gsub("^S$", "PDO.ASTURIAS", .))
t2$CA <- fct_relevel(t2$CA, sort)
```


```{r}
table(t2$years, t2$predoc)
table(t2$NIF)
table(t2$CA)
table(t2$area, t2$subarea)
tt <- table(t2$entity, t2$center)
```

```{r plots}
library("ggplot2")
  geom_count(aes(NIF, CA))
ggplot(t2) +
  geom_count(aes(entity, center))
ggplot(t2) +
  geom_bar(aes(entity, fill = area)) +
  facet_wrap(~CA) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
table(t2$CA)
library("dplyr")
t2 %>% 
  group_by(NIF, center) %>% 
  summarise(people = sum(predoc)) %>% 
  filter(people != 0) %>% 
  arrange(desc(people))
t2 %>% ggplot()+ 
  geom_bar(aes(CA, fill = area)) + 
  scale_fill_viridis_d() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Retos investigación 2018", x = "Comunidades autónomas", 
       y = "Proyectos")
t2 %>% 
  group_by(center, NIF) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  mutate(Site = paste(center, NIF)) %>% 
  head(n = 25) %>% 
  ggplot() + 
  geom_col(aes(fct_reorder(Site, order(n, decreasing = TRUE)), n))  +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank()) +
  labs(title = "Retos investigación 2018", x = "Centros", 
       y = "Proyectos concedidos")

```

## Economic
```{r}
anex2 <- anexes_field[[2]]
anex2t <- sapply(anex2, function(page){
  page[lengths(page) == 12]
})
anex2t2 <- lapply(anex2t[lengths(anex2t) > 1], function(page){
  t(simplify2array(page))
})
t2b <- do.call(rbind, anex2t2)
colnames(t2b) <- c("reference", "total", "direct", "indirect", "reimbursable", "grant", 
                  "grant_2019", "reimbursable_2019", "total_2019",
                  "grant_2020", "grant_2021", "grant_2022")
t2b <- apply(t2b, 2, function(x){
  y <- gsub("\\.", "", x)
  gsub(",", ".", y)
})
t2b <- as.data.frame(t2b)
t2b[, 2:12] <- sapply(t2b[, 2:12], fact2_number)
```
```{r}
library("ggforce")
# With help from: https://stackoverflow.com/a/32265122/2886003
euro_format <- dollar_format(suffix = "€", prefix = "")
ggplot(t2b) +
  geom_histogram(aes(total), bins = 100) +
  theme_bw() +
  scale_x_continuous(labels = euro_format) +
  facet_zoom(xlim = c(0, 400000)) +
  labs(y = "Projects", x = "Amount per project", title = "Money per project")

ggplot(t2b) +
  geom_point(aes(total, direct))
ggplot(t2b) +
  geom_point(aes(total, indirect))
ggplot(t2b) +
  geom_point(aes(grant, reimbursable)) +
  scale_x_continuous(labels = euro_format) +
  scale_y_continuous(labels = euro_format) +
  labs(x = "Subvención", y = "Reembolsable", title = "Relación financiera de los proyectos") + 
  theme_bw()
```


```{r}
df <- merge(t2b, t2)
ggplot(df) +
  geom_point(aes(grant, reimbursable, col = as.factor(years), 
                 size = as.factor(years))) +
  scale_x_continuous(labels = euro_format) +
  scale_y_continuous(labels = euro_format) +
  labs(x = "Subvención", y = "Reembolsable", 
       title = "Relación financiera de los proyectos",
       col = "Años", size = "Años") + 
  theme_bw()

df %>% 
  group_by(CA) %>% 
  summarise(total_CA =  sum(total), n = n(), mean = total_CA/n, sd = sd(total)) %>% 
  ggplot() + 
  geom_point(aes(total_CA, n, col = CA), size = 6) +
  theme_bw() +
  scale_x_continuous(labels = ) +
  labs(x = "Total €", y = "Projects", title = "Projects", col = "Comunidad autómoma")

df %>% 
  ggplot() +
  geom_violin(aes(CA, total, col = CA)) +
  geom_jitter(aes(CA, total, col = CA)) +
  theme_bw() +
  scale_y_continuous(labels = euro_format) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Euros per project", title = "Projects and CA", 
       x = "Comunidad Autónoma")

df %>% 
  group_by(CA, area) %>% 
  count() %>% 
  ggplot() +
  geom_point(aes(CA, area, size = n)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.ontop = element_blank()) +
  labs(y = "Projects", title = "Projects and CA", 
       x = "Comunidad Autónoma")

```


## Rechazados

```{r}
anex4 <- anexes_field[[3]]
anex4t <- sapply(anex4, function(page){
  page[lengths(page) == 7]
})
anex4t2 <- lapply(anex4t[lengths(anex4t) > 1], function(page){
  t(simplify2array(page))
})
t4b <- do.call(rbind, anex4t2)
colnames(t4b) <- c("reference", "area", "subarea", "entity", "center", "NIF", "CA")
t4b <- as.data.frame(t4b)
t4b$CA <- fct_relevel(t4b$CA, sort)
t4b %>% 
  group_by(CA) %>% 
  count() %>% 
  ggplot() +
  geom_col(aes(CA, n)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## All

```{r}
df2 <- merge(df, t4b, all = TRUE)
df2$CA <- fct_relevel(df2$CA, sort)
df2$NIF <- fct_relevel(df2$NIF, sort)
numeric_colnames <- colnames(df)
numeric_colnames <- numeric_colnames[-c(1, 13:18)]
df2[, numeric_colnames][is.na(df2[, numeric_colnames])] <- 0
```


```{r}
df2 %>% 
  group_by(NIF) %>% 
  summarise(n = n(), granted = sum(total != 0), percent = granted/n*100) %>% 
  arrange(desc(percent), desc(n)) %>% 
  ggplot() +
  geom_point(aes(n, granted, col = percent)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme_bw()

df2 %>% 
  group_by(area) %>% 
  summarise(n = n(), granted = sum(total != 0), percent = granted/n*100) %>% 
  arrange(desc(percent), desc(n)) %>% 
  ggplot() +
  geom_point(aes(n, granted, col = percent)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  theme_bw()
```

